---
title: "History of US Congress"
author: "Simon Garnier"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    css: custom.css
    fig_caption: true
    fig_width: 8
    number_sections: true
    theme: cosmo
    toc: true
    csl: plos.csl
    bibliography: library.bib
---

--- 

# Setup

```{r, 'setup', message=FALSE}
# Load libraries
library(yaml)
library(lubridate)
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(graphZoo)
library(scales)

# Create pretty, color blind friendly palette
cbf <- brewer.pal(8, "Dark2")

# Download raw data
if (!file.exists("../data/legislators-current.yaml")) {  
  download.file("https://raw.githubusercontent.com/unitedstates/congress-legislators/master/legislators-current.yaml",
                destfile = "../data/legislators-current.yaml", method = "wget")
}

if (!file.exists("../data/legislators-historical.yaml")) {  
  download.file("https://raw.githubusercontent.com/unitedstates/congress-legislators/master/legislators-historical.yaml",
                destfile = "../data/legislators-historical.yaml", method = "wget")
}
```

[Back to top]("#")

---

# Prepare data

```{r, 'data'}
# Function to help convert yaml database to data.frame
convert_yaml <- function(l) {
  data.frame(id = l$id$bioguide,
             name_first = l$name$first,
             name_last = l$name$last,
             birthday = as.Date(ifelse(is.null(l$bio$birthday), NA, l$bio$birthday)),
             gender = l$bio$gender,
             religion = ifelse(is.null(l$bio$religion), "Unknown", l$bio$religion),
             type = sapply(l$terms, function(x) { x$type }),
             start = as.Date(sapply(l$terms, function(x) { x$start })),
             end = as.Date(sapply(l$terms, function(x) { x$end })),
             state = sapply(l$terms, function(x) { x$state }),
             party = sapply(l$terms, function(x) { ifelse(is.null(x$party), NA, x$party) }),
             truc = 0)
}

# Load data
if (file.exists("../data/legislators.csv")) {
  dat <- read.csv("../data/legislators.csv",
                  colClasses = c(birthday = "Date", start = "Date", end = "Date"))
} else {
  historical <- do.call(rbind, lapply(yaml.load_file("../data/legislators-historical.yaml"), convert_yaml))
  current <- do.call(rbind, lapply(yaml.load_file("../data/legislators-current.yaml"), convert_yaml))
  dat <- rbind(historical, current)
  write.csv(dat, "../data/legislators.csv", row.names = FALSE)
}
```

[Back to top]("#")

---

# Compute statistics for each year

For someone to be considered a senator or representative in a given year, his/her term must start, end or be ongoing in any day of that year. Note that this approximation might result in a slightly higher number of senators and representatives than the congress is normally composed of in any given year (respectively 100 and 535 currently, but it changed over time). 

```{r, 'by_year', cache=TRUE}
years <- 1789:2014
stat <- {}

for (i in 1:length(years)) {
  sen <- filter(dat, (year(start) <= years[i] & year(end) >= years[i]) & type == "sen") %>%
    do(data.frame(number = c(as.vector(table(.$gender)), as.vector(table(.$religion))),
                  denomination = c(names(table(.$gender)), names(table(.$religion))))) %>%
    mutate(type = c("gender", "gender", rep("religion", length(number) - 2)),
           house = "sen",
           year = years[i])
  
  rep <- filter(dat, (year(start) <= years[i] & year(end) >= years[i]) & type == "rep") %>%
    do(data.frame(number = c(as.vector(table(.$gender)), as.vector(table(.$religion))),
                  denomination = c(names(table(.$gender)), names(table(.$religion))))) %>%
    mutate(type = c("gender", "gender", rep("religion", length(number) - 2)),
           house = "rep",
           year = years[i])
  
  stat <- rbind(stat, rep, sen)
}
```

[Back to top]("#")

---

# Percent women per year

```{r, 'parity', dpi=300, fig.height=6, fig.width=8}
parity <- filter(stat, type == "gender" & year >= 1866) %>%
  group_by(year, house) %>%
  summarize(prop_women = number[denomination == "F"] / sum(number)) %>%
  ungroup()

g <- ggplot(parity, aes(x = year, y = prop_women, color = house)) + 
  geom_hline(yintercept = 0.5, linetype = 2) +
  geom_hline(yintercept = 0, size = 0.125) + 
  geom_line(size = 1.5) +  
  geom_segment(data = data.frame(x = c(1866, 1917, 1922), y = 0.07, yend = 0.015), 
               aes(x = x, xend = x, y = y, yend = yend),
               arrow = arrow(length = unit(2, "mm")), color = "black") +
  annotate("text", x = c(1866, 1917, 1922), y = 0.075,
           label = c("1866: 1st woman to run for US Congress", 
                     "1917: 1st woman in US House", 
                     "1922: 1st woman in US Senate"),
           family = "Avenir Next", hjust = 0, vjust = 0.5, size = 4, angle = 90) +
  annotate("text", x = 2010, y = 0.495, label = "Parity", 
           family = "Avenir Next", hjust = 0, vjust = 1, size = 5) +
  scale_color_manual(values = cbf[1:2], name = "", labels = c("House of Representatives   ", "Senate"), 
                     guide = guide_legend(nrow = 1)) +
  scale_y_continuous(labels = percent) + 
  xlab("Year") + ylab(NULL) +
  theme_graphzoo(base_size = 15, family = "Avenir Next") + 
  theme(legend.position = c(-0.089, 1.175), legend.justification = c(0, 1),
        panel.grid.minor = element_blank())

g <- addTitle(g, "Percentage of women in US Congress", n.lines = 2)
 
g <- addBanner(g, font.size = 4, heights = c(1, 0.05 * 8 / 6),
               l.txt = "GRAPHZOO.TUMBLR.COM", 
               r.txt = "SOURCE: THEUNITEDSTATES.IO")

g
```


<p class="caption">**Figure 1:** Figure caption.</p>

[Back to top]("#")

---

